The impact of gas bulk rotation on the Lyman $\alpha$ line

### 2014-04-24 paper/ms.tex
new abstract

We present results of radiative transfer calculations to measure the
impact of gas bulk rotation on the morphology of the Lyman $\alpha$
emission line in distant galaxies. We model a galaxy as a sphere with
a homogeneous mixture of dust and hydrogen at a constant
temperature. These spheres have a solid-body rotation with maximum
velocities in the range $0-300$ \kms and neutral hydrogen optical
depths in the range $\tau_{\rm H}=10^{5}-10^{7}$. We consider cases 
of a single central \lya source, and a uniform distribution of sources. 
Our main finding are that the escape fraction of \lya photons, and
the intensity at the line center increases with rotational velocity. 
For homogeneously distributed sources the line becomes single peaked 
at rotational velocities larger than half the line width in the static
case. We also find that this anisotripic velocity configuration does
not induce any significant anisotropy in the spectral properties.
For radiation sources located off-center, the line morphology
presents  a range of single, double and triple peaks. We show how
these results  are useful to interpret recent spectroscopic results of distant 
$z\sim 2-3$ star forming galaxies.

\end{abstract}

>>>>

We present results of radiative transfer calculations to measure the
impact of gas bulk rotation on the morphology of the Lyman $\alpha$
emission line in distant galaxies. We model a galaxy as a sphere with
an homogeneous mixture of dust and hydrogen at a constant
temperature. These spheres have a solid-body rotation with maximum
velocities in the range $0-300$ \kms and neutral hydrogen optical
depths in the range $\tau_{\rm H}=10^{5}-10^{7}$. We consider cases 
of a single central \lya source, and a uniform distribution of sources. 
Our main result is that rotation introduces a dependence of the
line morphology with  viewing angle. Observations parallel to the
rotation axis yield lines similar to the static case. The greatest
difference is observed perpendicular to the rotation axis where the
intensity at the line center increases with rotational velocity.  For
homogeneously distributed sources the line becomes single peaked  at
rotational velocities larger than half the line width in the static
case. However, rotation does not not induce any significant anisotropy
in the total line intensity or in the escape fraction. 
\end{abstract}
